---
title: "Health and Economic Impact of Storm Damage"
author: "Brian L. Fuller"
date: "October 20, 2015"
output: 
  html_document: 
    keep_md: yes
---

Synopsis
-----

Your data analysis must address the following questions:

1. Across the United States, which types of events (as indicated in the EVTYPE
variable) are most harmful with respect to population health?
2. Across the United States, which types of events have the greatest economic
consequences?



Data Processing
-----

I'll begin by downloading the bz2 file if necessary and reading directly from
it. Based on the questions to be answered, I will not need all of the columns. 
I will specifically need event type (EVTYPE) and the columns which denote
population health (23 FATALITIES, 24 INJURIES) and economic consequence (25 PROPDMG,
26 PROPDMGEXP, 27 CROPDMG, and 28 CROPDMGEXP). I will also include several other
columns (7 STATE, 5 COUNTY, 6 COUNTYNAME, 2 BGN_DATE) to aid with analysis.

```{r, ReadRawData, cache=TRUE}
library(readr)

URL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
DestFile = "repdata_data_StormData.csv.bz2"

# download bz2 version of input data file if needed, and unzip
if(!file.exists(DestFile)){
     library(R.utils) 
     dlMethod <- "curl"
     if(substr(Sys.getenv("OS"),1,7) == "Windows") dlMethod <- "wininet"
     download.file(URL,destfile = DestFile, method = dlMethod, mode = "wb")
     ##bunzip2(DestFile, "StormData.csv")
}
myData <- read_csv(DestFile,
                   n_max = 902298, col_names = TRUE,
                   col_types = "_c__nccc______________nnncnc_______c_")
```

In order to make the date column (BGN_DATE) useful, convert character date
representation to an actual date representation.
```{r}
myData$BgnDate <- as.Date(myData$BGN_DATE, "%m/%d/%Y")
```

This is what our slightly conditioned dataset looks like
```{r}
head(myData)
```

Results
-----






